Tracing the cultural evolution of meaning at the semantics-pragmatics interface

### 2017-03-20 paper/paper-v3.tex
revision

A particularly productive and well studied class of systematic pragmatic enrichments are scalar
implicatures
\citep{horn:1984,Hirschberg1985:A-Theory-of-Sca,LevinsonPragmatics1983,Geurts2010:Quantity-Implic}. Usually,
the utterance of a sentence like ``I own some of Johnny Cash's albums'' will be taken to mean
that the speaker does not own all of them. This is because, if the speaker had them all, he
could have used the stronger word \emph{all} instead of \emph{some} in his utterance, 
thereby making a more informative statement. Scalar implicatures, especially the
inference from \emph{some} to \emph{some but not all}, have been studied extensively, both
theoretically
\citep[e.g.,][]{Sauerland2004:Scalar-Implicat,ChierchiaFox2008:The-Grammatical,Rooyvan-RooijJagerde-Jager2012:Explaining-Quan}
as well as experimentally
\citep[e.g.,][]{BottNoveck2004:Some-Utterances,huang+snedeker:2009,GrodnerKlein2010:Some-and-Possib,GoodmanStuhlmuller2013:Knowledge-and-I,DegenTanenhaus2012:Processing-Scal}. While
there is much dispute in this domain about many details, a position endorsed by a
clear majority is that a word like \emph{some} is underspecified to mean \emph{some and maybe
  all} and that the enrichment to \emph{some but not all} is part of some regular 
process with roots in pragmatics.

@@

An elegant formal approach to capture the interaction between expressivity and learnability
is the \emph{replicator mutator dynamic}
\citep{Hofbauer1985:The-Selection-M,nowak+etal:2000,NowakKomarova2001:Evolution-of-Un,hofbauer+sigmund:2003,Nowak2006:Evolutionary-Dy}. In
its simplest, discrete-time formulation, the RMD defines the frequency $x'_i$ of each type $i$
in a population at the next time step as a function of: (i) the frequency $x_i$ of each type
$i$ before the step, (ii) the fitness $f_i$ of each type $i$ before the step, and (iii) the
probability $Q_{ji}$ that an agent who wants to imitate, adopt, or learn the type of an agent
with type $j$ ends up acquiring type $i$:
\begin{align}
  \label{eq:RMD_discrete}
  x'_i = \sum_j Q_{ji} \frac{x_jf_j}{\sum_k x_k f_k}\,.
\end{align}
The RMD consists of two components: fitness-based selection and transmission biases. This
becomes most transparent when we consider an equivalent formulation in terms of a step-wise
application of the discrete-time replicator dynamic \citep{TaylorJonker1978:Evolutionary-St} on the initial population vector $\vec{x}$
and its subsequent multiplication with a mutation matrix $Q$:

\begin{align}
  \label{eq:RMD_discrete_recast}
  x'_i & = (\text{M}(\text{RD}(\vec{x})))_i\,,
\end{align}
where
\begin{align*}
      \left ( \text{RD}(\vec{x}) \right )_i 
         = \frac{x_i f_i}{\sum_k x_k f_k}
 \ \ \ \ \text{and} \ \ \ \ 
  (\text{M}(\vec{x}))_i = (\vec{x} \cdot Q)_i = \left ( \sum_j
          x_j Q_{ji} \right)_i\,.
\end{align*}
If the transmission matrix $Q$ is trivial in the sense that $Q_{ji}=1$ whenever $j=i$, the
dynamic reduces to the replicator dynamic. The replicator dynamic is a model of fitness-based
selection in which the relative frequency of type $i$ will increase with a gradient
proportional to its average fitness in the population. This dynamic is popular and
versatile because it can be derived from many abstract processes of biological and cultural
transmission and selection \citep[for overview and several derivations
see][]{Sandholm2010:Population-Game}, including conditional imitation
\citep[e.g.,][]{Helbing1996:A-Stochastic-Be,Schlag1998:Why-Imitate-and} or reinforcement
learning \citep[e.g.,][]{BorgersSarin997:Learning-Throug,Beggs2005:On-the-Converge}. If fitness
$f_i$ is the same for all types $i$, the replicator step is the identity map
$ \left ( \text{RD}(\vec{x}) \right )_i = x_i$ and the dynamic reduces to a process of
iteration of the transmission bias encoded in $Q$. In this way, the process in
(\ref{eq:RMD_discrete}), equivalently (\ref{eq:RMD_discrete_recast}), contains a model of
iterated learning \citep{griffiths+kalish:2007}. \mf{should we include a simple example here? I 
have an example from a lecture ready at hand; it's a simple coordination game in a
one-population setting.} \tb{Yes, I think that's a good idea. We can always cut it out again if the reviewers prefer more compression but let's explain more and cut later. The paper is relatively short at the moment.}

@@

We distinguish between two kinds of pragmatic behavior. {\em Literal interlocutors} produce and
interpret messages literally, being guided only by their lexica. {\em Pragmatic interlocutors}
instead engage in mutual reasoning to inform their choices. Recent probabilistic models of
rational language use
\citep{franke:2009,frank+goodman:2012,FrankeJager2015:Probabilistic-p,GoodmanFrank2016:Pragmatic-Langu}
capture different types of pragmatic behavior in a reasoning hierarchy. The hierarchy's bottom,
level $0$, corresponds to literal language use. Pragmatic language users of level $n + 1$
act (approximately) rational with respect to level-$n$ behavior of their
interlocutors. (\ref{h:level0}) and (\ref{s:level0}) define probabilistic behavior of literal
hearers and speakers respectively: \mf{is it a problem that $S$ denotes speakers and the set of
states?} \tb{I think that's fine. $S_n(\cdot \mid \cdot)$ is not used beyond this section. Otherwise we can, e.g., relabel $M$ as $F$(orms) and use $M$(eanings) instead of $S$(tates). I'm fine with either option.}
\begin{flalign}
&H_{0}(s \mid m;L) \propto pr(s) L_{sm} \label{h:level0}\\
&S_{0}(m \mid s;L) \propto \exp(\lambda \; L_{sm}) \label{s:level0}
\end{flalign}

According to (\ref{h:level0}), a literal hearer's interpretation of a message $m$ as a state
$s$ depends on her lexicon and her prior over states, $pr \in \Delta(S)$. For simplicity, in
the following this prior is assumed to be uniform. A literal interpreter with lexicon $\Lbound$
from above would assign $\ssome$ a probability of $H_0(\ssome \mid \msome; \Lbound) = 1$ after
hearing $\msome$, while a literal interpreter with lexicon $\Llack$ would assign $\ssome$
probability $H_0(\ssome \mid \msome; \Llack) = 0.5$. As usual in probabilistic pragmatics models, speaker behavior is regulated by a soft-max
parameter $\lambda$, $\lambda \geq 1$ \citep{luce:1959,sutton+barto:1998}. As $\lambda$
increases, choices made in production are more rational in that higher values lead to behavior
that is increasingly in line with expected utility maximization. Expected utility of a message
$\messg$ in state $\state$ for a level $n+1$ speaker is here defined as $H_{n}(s|m;L)$, the probability that the hearer will assign to or choose the correct meaning. Put differently, how likely the hearer is thought to infer $\state$ when sent $\messg$. The formulation
in (\ref{s:level0}) also allows false messages to be sent with a small positive probability in
analogy to the definition of pragmatic speakers in probabilistic pragmatic models. This is also necessary to guarantee a mutation matrix with only positive entries (see below). If $\lambda = 1$ a literal
speaker with either lexicon $\Lbound$ or $\Llack$ produces $\msome$ in $\ssome$ with probability $S_{0}(\msome \mid
\ssome;L_{\text{bound},\text{lack}}) \approx .73$. The probability of producing $\msome$ in $\sall$ is $S_{0}(\msome \mid
\sall;\Lbound) \approx .27$ for $\Lbound$ and $S_{0}(\msome \mid
\sall;\Llack) = .5$ for $\Llack$. By contrast, if $\lambda = 20$, a literal speaker with either lexicon produces $\msome$ in $\ssome$ with probability $S_{0}(\msome \mid
\ssome;L_{\text{bound},\text{lack}}) \approx 1$. But the probability of producing $\msome$ in $\sall$ is $S_{0}(\msome \mid
\sall;\Lbound) \approx 1$ for $\Lbound$ whereas it is still $S_{0}(\msome \mid \sall;\Llack) = .5$ for $\Llack$ because it semantically associates $\msome$ with both states.

@@

Particularly important to our purpose is the fact that players of a level greater than $0$ using $\Llack$ can {\em pragmatically} associate $\msome$ with $\ssome$ over $\sall$,  in contrast to their literal counterparts. Intuitively, this is so because hearers reason a speaker to convey $\sall$ with $\mall$, as this has a higher chance of leading to mutual understanding than using underspecified $\msome$. Therefore they will expect $\msome$ to be more strongly associated with $\ssome$ than with $\sall$ as it is already possible to clearly convey the latter state with a different message. The converse reasoning pattern holds for the speaker. In this way, two distinct lexical conventions such as $\Lbound$ and $\Llack$ can give rise to similar observable linguistic behavior.

@@

In biological evolution, where types are expressed genetically, transmission infidelity comes
into the picture through infrequent and mostly random genetic mutations. However, an agent's lexicon
and pragmatic reasoning behavior is not inherited genetically. They need to be learned from
observation. Concretely, when agents of type $j$ want to adopt or imitate the linguistic
behavior of type $i$, they observe the linguistic behavior of type $i$ and need to infer what
their type is from that. Iterated learning is a process in which languages are learned repeatedly from the observation
of linguistic behavior of agents who have acquired the language from observation and inference
before as well. In the simplest case there is a single teacher and a single
learner. After sufficient training the learner becomes a teacher and produces behavior that
serves as input for a new learner. Due to the pressure towards learnability it exerts, iterated
learning alone generally leads to simpler and more regular languages (see \citealt{kirby+etal:2014}
and \citealt{tamariz+kirby:2016} for recent surveys).

@@

Expressivity and learnability are central to the cultural evolution of language. These
components can be modelled, respectively, as replication based on a measure of fitness in terms
of communicative efficiency and iterated Bayesian learning. Their interaction is described by
the discrete time replicator mutator dynamics in (\ref{eq:RMD_discrete}), repeated here:
\begin{align*}
  x'_i = \sum_j Q_{ji} \frac{x_jf_j}{\sum_k x_k f_k}\,.
\end{align*}
This equation defines the frequency $x'_i$ of type $i$ at the next time step, based on its
frequency $x_i$ before the step, its fitness $f_i$ and the probability that a learner infers
$i$ when observing the behavior of a type-$j$ agent. Fitness-based selection can be thought of
as biological (fitness as expected relative number of offspring) or cultural (fitness as 
likelihood of being imitated or repeated). The types that the dynamic operates on are pairs
consisting of a lexicon and a pragmatic use pattern. A type's expressivity depends on its
communicative efficiency within a population while its learnability depends on the fidelity by
which it is inferred by new generations of learners. The learners' task is consequently to
perform a joint inference over types of linguistic behavior and lexical meaning.

@@

We begin setting the stage by defining the learning prior and the set of types under consideration. Drawing from our preceding discussion, functional pressure on successful communication combined with learning pressures in the form of a bias that favors simpler semantics may lead to the selection of $\Llack$-like semantics and to the majority view of a division of labor between semantics and pragmatics. We first showcase the effects that expressivity and learnability have in isolation before turning to those resulting from their combination.

@@

where $Compl(c)$ is the complexity of the minimal derivation length of $c$ according to the grammar. There are many ways to define priors over lexica (see, e.g., \citealt{goodman+etal:2008, piantadosi+etal:2012a}) but the key assumption here, common to all of them, is that simple representational expressions should be favored over more complex ones. We should stress that these details -- from the generative grammar to its complexity measure -- are to be regarded as an instantiation of our general assumptions instead of as empirical claims in their own right.

\subsubsection{Lexica, signaling behavior \& types}
We consider a state space with three states $\States = \set{\snone, \ssome, \sall}$. \tb{Should we say something about the inclusion of $\snone$? It may be more natural to introduce this state earlier but may unnecessarily complicate the lexica showcased in the previous section.} This space can be
thought of as a partition of possible worlds into cells where none, some or all of the $A$s are
$B$s, for some arbitrary fixed predicates $A$ and $B$. Eight concepts can be distinguished
based on their truth or falsity in three world states, six of which are not contradictory
 or tautological. These are listed with mnemonic names in
Table~\ref{tab:concepts} together with their complexity according to the grammar given in Table \ref{tab:grammar}.

@@

A lexicon $L$ is a mapping $\Messgs \rightarrow C$ from messages to concepts. With three
messages there are $6^3 = 216$ possible lexica. Some assign the same concept to more than one message and others lexicalize the same meanings but associate them with different messages. The main results reported in the following do not hinge on the inclusion of these lexica but such an unaltered space does better illustrate the influence and differences of the linguistic pressures we consider. Additionally, the competition of different languages that lexicalize the same concepts through different forms allows us to take serious the problem of only one being selected for even though multiple alternatives of equal expressivity {\em and} learnability exist.

Out of these possible lexica, three kinds are of particular relevance. First, lexica that assign the same concept to more than one message. Such lexica lack in expressivity but are favored by the learning bias should they lexicalize simple concepts. Second, lexica that conventionalize upper-bounds to realize a (quasi-)partition of the state space. As discussed before in relation to $\Lbound$, users of these lexica need not resort to pragmatic reasoning to convey an upper-bound with weak scalar expressions. Instead, this information is already provided by the semantics of their lexicon, giving them an advantage in terms of expressivity. Types with such lexica are expected to be the main contenders of the natural language like semantics that do not lexicalize an upper-bound for weak scalar expressions. Third and lastly, $\Llack$-style lexica that, paired with mutual reasoning, can convey an upper-bound pragmatically. There are six lexica of the second kind and six of the third. The following three lexica exemplify each kind, with $\Lall$ being a lexicon of the first kind. This lexicon is not optimal for communication as it assigns all its messages to the same state, $\sall$. This inevitably leads to a communicative disadvantage in its use. However, it may have a learnability advantage over our target $\Llack$ and its competitor $\Lbound$ in virtue of being the simplest lexicon, making it a priori favored by learners.

>>>>

A particularly productive and well studied class of systematic pragmatic enrichments are scalar
implicatures
\citep{horn:1984,Hirschberg1985:A-Theory-of-Sca,LevinsonPragmatics1983,Geurts2010:Quantity-Implic}. Usually,
the utterance of a sentence like ``I own some of Johnny Cash's albums'' will be taken to mean
that the speaker does not own all of them. This is because, if the speaker had them all, he
could have used the word \emph{all} instead of \emph{some} in his utterance, 
thereby making a more informative statement. Scalar implicatures, especially the
inference from \emph{some} to \emph{some but not all}, have been studied extensively, both
theoretically
\citep[e.g.,][]{Sauerland2004:Scalar-Implicat,ChierchiaFox2008:The-Grammatical,Rooyvan-RooijJagerde-Jager2012:Explaining-Quan}
as well as experimentally
\citep[e.g.,][]{BottNoveck2004:Some-Utterances,huang+snedeker:2009,GrodnerKlein2010:Some-and-Possib,GoodmanStuhlmuller2013:Knowledge-and-I,DegenTanenhaus2012:Processing-Scal}. While
there is much dispute in this domain about many details, a position endorsed by a
clear majority is that a word like \emph{some} is underspecified to mean \emph{some and maybe
  all} and that the enrichment to \emph{some but not all} is part of some regular 
process with roots in pragmatics.

@@

An elegant formal approach to capture the interaction between expressivity and learnability
is the \emph{replicator mutator dynamic}
\citep{Hofbauer1985:The-Selection-M,nowak+etal:2000,NowakKomarova2001:Evolution-of-Un,hofbauer+sigmund:2003,Nowak2006:Evolutionary-Dy}. In
its simplest, discrete-time formulation, the RMD defines the frequency $x'_i$ of each type $i$
in a population at the next time step as a function of: (i) the frequency $x_i$ of each type
$i$ before the step, (ii) the fitness $f_i$ of each type $i$ before the step, and (iii) the
probability $Q_{ji}$ that an agent who wants to imitate, adopt, or learn the type of an agent
with type $j$ ends up acquiring type $i$:
\begin{align}
  \label{eq:RMD_discrete}
  x'_i = \sum_j Q_{ji} \frac{x_jf_j}{\sum_k x_k f_k}\,.
\end{align}
The RMD consists of two components: fitness-based selection and transmission biases. This
becomes most transparent when we consider an equivalent formulation in terms of a step-wise
application of the discrete-time replicator dynamic \citep{TaylorJonker1978:Evolutionary-St} on the initial population vector $\vec{x}$
and its subsequent multiplication with a mutation matrix $Q$:
\begin{align}
  \label{eq:RMD_discrete_recast}
  x'_i & = (\text{M}(\text{RD}(\vec{x})))_i\,,
\end{align}
where
\begin{align*}
      \left ( \text{RD}(\vec{x}) \right )_i 
         = \frac{x_i f_i}{\sum_k x_k f_k}
 \ \ \ \ \text{and} \ \ \ \ 
  (\text{M}(\vec{x}))_i = (\vec{x} \cdot Q)_i = \left ( \sum_j
          x_j Q_{ji} \right)_i\,.
\end{align*}
If the transmission matrix $Q$ is trivial in the sense that $Q_{ji}=1$ whenever $j=i$, the
dynamic reduces to the replicator dynamic. The replicator dynamic is a model of fitness-based
selection in which the relative frequency of type $i$ will increase with a gradient
proportional to its average fitness in the population. This dynamic is popular and
versatile because it can be derived from many abstract processes of biological and cultural
transmission and selection \citep[for overview and several derivations
see][]{Sandholm2010:Population-Game}, including conditional imitation
\citep[e.g.,][]{Helbing1996:A-Stochastic-Be,Schlag1998:Why-Imitate-and} or reinforcement
learning \citep[e.g.,][]{BorgersSarin997:Learning-Throug,Beggs2005:On-the-Converge}. If fitness
$f_i$ is the same for all types $i$, the replicator step is the identity map
$ \left ( \text{RD}(\vec{x}) \right )_i = x_i$ and the dynamic reduces to a process of
iteration of the transmission bias encoded in $Q$. In this way, the process in
(\ref{eq:RMD_discrete}), equivalently (\ref{eq:RMD_discrete_recast}), contains a model of
iterated learning \citep{griffiths+kalish:2007}. \mf{should we include a simple example here? I 
have an example from a lecture ready at hand; it's a simple coordination game in a
one-population setting.} \tb{Yes, I think that's a good idea. We can always cut it out again if the reviewers prefer more compression but let's explain more and cut later. The paper is relatively short at the moment.}

@@

We distinguish between two kinds of linguistic behavior. {\em Literal interlocutors} produce and
interpret messages literally, being guided only by their lexica. {\em Pragmatic interlocutors}
instead engage in mutual reasoning to inform their choices. Recent probabilistic models of
rational language use
\citep{franke:2009,frank+goodman:2012,FrankeJager2015:Probabilistic-p,GoodmanFrank2016:Pragmatic-Langu}
capture different types of pragmatic behavior in a reasoning hierarchy. The hierarchy's bottom,
level $0$, corresponds to literal language use. Pragmatic language users of level $n + 1$
act (approximately) rational with respect to level-$n$ behavior of their
interlocutors. (\ref{h:level0}) and (\ref{s:level0}) define probabilistic behavior of literal
hearers and speakers respectively: \mf{is it a problem that $S$ denotes speakers and the set of
states?} \tb{I think that's fine. $S_n(\cdot \mid \cdot)$ is not used beyond this section. Otherwise we can, e.g., relabel $M$ as $F$(orms) and use $M$(eanings) instead of $S$(tates). I'm fine with either option.}
\begin{flalign}
&H_{0}(s \mid m;L) \propto pr(s) L_{sm} \label{h:level0}\\
&S_{0}(m \mid s;L) \propto \exp(\lambda \; L_{sm}) \label{s:level0}
\end{flalign}

According to (\ref{h:level0}), a literal hearer's interpretation of a message $m$ as a state
$s$ depends on her lexicon and her prior over states, $pr \in \Delta(S)$. For simplicity, in
the following this prior is assumed to be uniform. A literal interpreter with lexicon $\Lbound$
from above would assign $\ssome$ a probability of $H_0(\ssome \mid \msome; \Lbound) = 1$ after
hearing $\msome$, while a literal interpreter with lexicon $\Llack$ would assign $\ssome$
probability $H_0(\ssome \mid \msome; \Llack) = 0.5$. As usual in probabilistic pragmatics models, speaker behavior is regulated by a soft-max
parameter $\lambda$, $\lambda \geq 1$ \citep{luce:1959,sutton+barto:1998}. As $\lambda$
increases, choices made in production are more rational in that they are increasingly in line with expected utility maximization. Expected utility of a message
$\messg$ in state $\state$ for a level $n+1$ speaker is here defined as $H_{n}(s|m;L)$, the probability that the hearer will assign to or choose the correct meaning. Put differently, expected utility indicates how likely the hearer is thought to infer $\state$ when sent $\messg$. The formulation
in (\ref{s:level0}) also allows false messages to be sent with a small positive probability in
analogy to the definition of pragmatic speakers in probabilistic pragmatic models. This is also necessary to guarantee a mutation matrix with only positive entries (see below). If $\lambda = 1$ a literal
speaker with either lexicon $\Lbound$ or $\Llack$ produces $\msome$ in $\ssome$ with probability $S_{0}(\msome \mid
\ssome;L_{\text{bound},\text{lack}}) \approx .73$. The probability of producing $\msome$ in $\sall$ is $S_{0}(\msome \mid
\sall;\Lbound) \approx .27$ for $\Lbound$ and $S_{0}(\msome \mid
\sall;\Llack) = .5$ for $\Llack$. By contrast, if $\lambda = 20$, a literal speaker with either lexicon produces $\msome$ in $\ssome$ with probability $S_{0}(\msome \mid
\ssome;L_{\text{bound},\text{lack}}) \approx 1$. But the probability of producing $\msome$ in $\sall$ is $S_{0}(\msome \mid
\sall;\Lbound) \approx 1$ for $\Lbound$ whereas it is still $S_{0}(\msome \mid \sall;\Llack) = .5$ for $\Llack$ because it semantically associates $\msome$ with both states.

@@

Particularly important to our purpose is the fact non-literal interlocutors using $\Llack$ can {\em pragmatically} associate $\msome$ with $\ssome$ over $\sall$,  in contrast to their literal counterparts. Intuitively, this is so because hearers reason a speaker to convey $\sall$ with $\mall$, as this message has a higher chance of leading to mutual understanding than using underspecified $\msome$. Therefore they will expect $\msome$ to be more strongly associated with $\ssome$ than with $\sall$ as it is already possible to clearly convey the latter state $\mall$. The converse reasoning pattern holds for the speaker. In this way, two distinct lexical conventions such as $\Lbound$ and $\Llack$ can give rise to similar observable linguistic behavior. In this case, associating $\msome$ with $\ssome$ and $\mall$ with $\sall$.

@@

In biological evolution, where types are expressed genetically, transmission infidelity comes
into the picture through infrequent and mostly random genetic mutations. However, an agent's lexicon
and pragmatic reasoning behavior is not inherited genetically. They need to be learned from
observation. Concretely, when agents of type $j$ want to adopt or imitate the linguistic
behavior of type $i$, they observe the overt linguistic behavior of type $i$ and need to infer what
their type is from that. Iterated learning is a process in which languages are learned repeatedly from the observation
of linguistic behavior of agents who have acquired the language from observation and inference
before as well. In the simplest case there is a single teacher and a single
learner. After sufficient training the learner becomes a teacher and produces behavior that
serves as input for a new learner. Due to the pressure towards learnability it exerts, iterated
learning alone generally leads to simpler and more regular languages (see \citealt{kirby+etal:2014}
and \citealt{tamariz+kirby:2016} for recent surveys).

@@

Expressivity and learnability are central to the cultural evolution of language. These
components can be modelled, respectively, as replication based on a measure of fitness in terms
of communicative efficiency and iterated Bayesian learning. Their interaction is described by
the discrete time replicator mutator dynamics in (\ref{eq:RMD_discrete}), repeated here:
\begin{align*}
  x'_i = \sum_j Q_{ji} \frac{x_jf_j}{\sum_k x_k f_k}\,.
\end{align*}
This equation defines the frequency $x'_i$ of type $i$ at the next time step, based on its
frequency $x_i$ before the step, its fitness $f_i$, and the probability that a learner infers
$i$ when observing the behavior of a type-$j$ agent. Fitness-based selection can be thought of
as biological (fitness as expected relative number of offspring) or cultural (fitness as 
likelihood of being imitated or repeated). The types that the dynamic operates on are pairs
consisting of a lexicon and a pragmatic use pattern. A type's expressivity depends on its
communicative efficiency within a population while its learnability depends on the fidelity by
which it is inferred by new generations of learners. The learners' task is consequently to
perform a joint inference over types of linguistic behavior and lexical meaning.

@@

We begin setting the stage by defining the learning prior and the set of types under consideration. Drawing from our preceding discussion, functional pressure on successful communication combined with learning pressures in the form of a bias that favors simpler semantics may lead to the selection of $\Llack$-like semantics and to a division of labor between semantics and pragmatics. We first showcase the effects that expressivity and learnability have in isolation before turning to those resulting from their combination.

@@

where $Compl(c)$ is the complexity of the minimal derivation length of $c$ according to the LOT-grammar. There are many ways to define priors over lexica (see, e.g., \citealt{goodman+etal:2008, piantadosi+etal:2012a}) but the key assumption here, common to all of them, is that simple representational expressions should be favored over more complex ones. We should stress that these details -- from the generative grammar to its complexity measure -- are to be regarded as an instantiation of our general assumptions instead of as empirical claims in their own right.

\subsubsection{Lexica, signaling behavior \& types}
We consider a state space with three states $\States = \set{\snone, \ssome, \sall}$. \tb{Should we say something about the inclusion of $\snone$? It may be more natural to introduce this state earlier but may unnecessarily complicate the exposition of signaling behavior in the previous section.} This space can be
thought of as a partition of possible worlds into cells where none, some or all of the $A$s are
$B$s, for some arbitrary fixed predicates $A$ and $B$. Eight concepts can be distinguished
based on their truth or falsity in three world states, six of which are not contradictory
 or tautological. These are listed with mnemonic names in
Table~\ref{tab:concepts} together with their complexity according to the grammar in Table \ref{tab:grammar}.

@@

A lexicon $L$ is a mapping $\Messgs \rightarrow C$ from messages to concepts. With three
messages there are $6^3 = 216$ possible lexica. Some assign the same concept to more than one message and others lexicalize the same concepts but associate them with different messages. The main results reported in the following do not hinge on the inclusion of these lexica but this unaltered space does better illustrate the influence and differences of the linguistic pressures we consider. Additionally, the competition of different languages that lexicalize the same concepts through different forms allows us to take serious the problem of only one being selected for even though multiple alternatives of equal expressivity {\em and} learnability exist.

Out of these possible lexica, three kinds are of particular relevance. First, lexica that assign the same concept to more than one message. Such lexica lack in expressivity but may be favored by the learning prior should they lexicalize simple concepts. Second, lexica that conventionalize upper-bounds to realize a (quasi-)partition of the state space. As discussed before in relation to $\Lbound$, users of these lexica need not resort to pragmatic reasoning to convey an upper-bound with weak scalar expressions. Instead, this information is already provided by the semantics of their lexicon, giving them an advantage in terms of expressivity. Types with such lexica are expected to be the main contenders of the natural language like semantics that do not lexicalize an upper-bound for weak scalar expressions. Third and lastly, $\Llack$-style lexica that, paired with mutual reasoning, can convey an upper-bound pragmatically. There are six lexica of the second kind and six of the third. The following three lexica exemplify each kind, with $\Lall$ being a lexicon of the first kind. This lexicon is not optimal for communication as it assigns all its messages to the same state, $\sall$. This inevitably leads to a communicative disadvantage in its use. However, it may have a learnability advantage over our target $\Llack$ and its competitor $\Lbound$ in virtue of being the simplest lexicon, making it a priori favored by learners.
